Bootstrap learning via modular concept discovery

### 2013-01-31 drafts/ijcai2013/ijcai13.tex
Update.
Changes.

\usepackage{ijcai13}
\usepackage{times}
\usepackage{amsmath} 
\usepackage{latexsym} 
\usepackage{ dsfont }
\usepackage{tikz}
\usepackage{circuitikz}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{alltt}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usetikzlibrary{arrows,automata, positioning, patterns}

\newtheorem{definition}{Definition}
\DeclareMathOperator*{\argmax}{arg\,max}

@@

\author{Eyal Dechter \\
MIT\\
USA \\
edechter@mit.edu
\And
Jon Malmaud \\
MIT \\ 
USA \\
malmaud@mit.edu
\And 
Ryan P.  Adams \\
Harvard University\\
USA \\
rpa@seas.harvard.edu
\And
Josh Tenenbaum \\
MIT\\
USA \\
jbt@mit.edu
}

@@

\section{Introduction}

Imagine you are trying to learn to bake via trial and error. You know
what kind of ingredients are used in baking (flour, water, butter,
sugar, yeast, baking soda, etc.) and you know a bit about how to
combine these ingredient (mix, whisk, etc.) and how to apply heat
(oven). But baking is a notoriously fragile process: ratios of
ingredients, the order of mixing, cooking temperatures, resting times,
all make a huge difference, and the outcome of the baking process is
mostly all or none. Either the outcome is what you expected or quite
unlike it. Even if you do create something similar to your intent, it
is difficult to assign blame for the outcome to any one step or
decision along the way.

Many difficult learning problems have this structure -- learning
language, developing technologies and scientific theories, learning to
play instruments or execute other complex motor patterns. These are
problems in which 1) there is a large set of similar tasks; 2) their
solutions are best described as programs; 3) the solution space is
not smooth with respect to the outcome; 4) and the outcome is generally
uninformative unless you are already pretty close to a solution.

Intuitively, the way to solve such problems is to stumble on solutions
to simple cases in the domain and extract conceptual building blocks
that can be composed to scale up to complex cases. To return to the
baking example, cookies are much simpler than cakes. There are fewer
steps to get just right. But once you have figured out a few simple
pastries, the more complex ones are mostly just variations and
combinations of techniques learned along the way.

Of course, knowledge is not acquired wholly by trial and error. But in
many domains, there is at least initially such a period of ignorance,
and the question we are trying to tackle here is how you ever get off
the ground.

To frame this learning paradigm computationally, we make two modeling
choices. First, to emphasize the impoverished reward signals in these
domains, we model the learning task as an all-or-none optimization
problem: either the learner gets some reinforcement or she does
not. Second, to emphasize the lack of locally smooth structure in the
space of possible solutions, we choose as our solution space the space
of functional programs. These two choices are in contrast to the
paradigm tackled by most of machine learning, in which a locally
informative objective function guides search in a space that is
somewhat smooth.

%% Let a \emph{task} be a function $t : \mathcal{L} \rightarrow \{0,
%% 1\}$, where $\mathcal{L}$ is the set of all expressions in our
%% representation language. We say that $e \in \mathcal{L}$ \emph{solves}
%% task $t$ if $t(e) = 1$. Our goal is to solve as many tasks as
%% possible, as efficiently as possible. That is, suppose that there is a
%% distribution over tasks $t \sim  p_t(\cdot)$, we would like an ordering $O
%% = e_1, \dots $ on expressions in $\mathcal{L}$ such that if $p_t(t_1)
%% > p_t(t_2)$ then the first expression in $O$ to solve $t_1$ comes
%% before the first expression in $O$ to solve $t_2$.

%% We motivate a solution to this problem by considering a heirarchical
%% Bayesian formulation of the task distribution. Suppose we believe that
%% there is some underlying structure to $p_t(\cdot)$ that is explained by
%% the fact that for each task $t$ there is a latent expressions $e \in
%% \mathcal{L}$ drawn from a stochastic grammar $\mathcal{G}$ and that

%% \begin{align}
%% e &\sim p_\mathcal{L}(\cdot) = \mathcal{G}\\
%% t | e &\sim \text{U}\{ \text{tasks solved by }e\}.
%% \end{align}

%% Then $p_t(t) = \sum_{e_j \in \mathcal{L} \text{ s.t. } t(e_j) = 1}
%% p_\mathcal{L}(e_j)$. If we make the assumption that for the set $E$ of
%% expressions in this sum, there is some expression $e^*$ such that
%% $p_\mathcal{L}(e^*) >> p_\mathcal{L}(e), e \neq e^*$, then we have, approximately, 
%% \[ p_t(t) \propto} p_\mathcal{L}(e^*). \]

%% This suggests a way to find the ordering $O$: infer a stochastic
%% grammar $G$, under the assumption that $G$ favors one solution to a
%% given task over any other, and order the elements of $\mathcal{L}$
%% according to $G$.

%% Thus, our basic procedure will be to search for a single ``good''
%% solution for as many tasks as we can, and to use those solutions to
%% estimate a stochastic grammar. Our search will be performed in the
%% order of the most recently inferred grammar. So as our ordering
%% becomes more efficient and we can solve more tasks, we learn better
%% stochastic grammars.

%% Formally then, our goal is to solve the following minimal feedback
%% multi-task program induction problem: suppose we are given set of
%% \emph{tasks} $T=\{t_k\}_{k=1}^K$ where each task is a function $t_i :
%% \mathcal{L} \rightarrow \{0, 1\}$, where $\mathcal{L}$ is the set of
%% programs in our language.  We say that $ p \in \mathcal{L}$
%% \emph{solves} $t_i$ if $ t_i\; p = 1$. Our goal is to find a set of
%% solutions $X = \{x_k\}_{k=1}^k \subset \mathcal{L}$ such that

%% \[
%% X = \argmax_{x_1, \dots, x_K \in L} \sum_k t_k(x_k).
%% \]

Based on the motivation set out above, we present an iterative
strategy. Given a stochastic grammar $\mathcal{G}$ which defines a
distribution over the primitives from $\mathcal{L}$ and a integer
\emph{frontier size} $N$, we perform two steps in each iteration.  In
the first step, we explore the $N$ highest probability programs
according to $\mathcal{G}$. In the second step, we compress the
programs that solve any of our tasks, and we use that compression to
define a new grammar. Repeating these two steps, we effectively
reparameterize our search space to include programs that are more likely
to be solutions to our tasks.

The remainder of this paper will include three major sections. First,
we will describe the representation language we chose and the goals of
the exploration and compressions steps. Following that, we will
discuss the algorithmic details that render our procedure
feasible. Finally, we present results from two experiments, involving
the domains of symbolic regression and boolean circuit learning.

@@

@@

Following~\cite{liang10programs} and~\cite{Briggs:2008}, we use
combinatory logic -- a variable-free subset of the polymorphic
simply-typed lambda calculus~\cite{Pierce_2002} -- as
our program representation language. In short, the combinatory logic
introduces a \emph{basis} of several primitive \emph{combinators} such
that any function in the lambda calculus can be alternatively written
as applications of the basis combinators. Some common basis
combinators are defined as follows:
\begin{align}
I\; x &\rightarrow  x \\
S\; f\; g\; x\; &\rightarrow (f\; x)\; (g\; x) \\ 
B\; f\; g\; x\; &\rightarrow (f\; x)\; g \\ 
C\; f\; g\; x\; &\rightarrow f\; (g\; x) \\ 
\end{align}

The basis
combinators are theoretically enough to express any Turing computable
function, but we will assume that our language has a variety of
primitives and, together with the basis combinators, we call these the
primitive combinators.

The basis combinators can themselves been expressed in the lambda
calculus. The lambda calculus has two basic operations -- application
and abstraction -- but in using the combinatory logic we sequester
uses of the abstraction operation inside the combinators, and thus
have a representation language that is effectively abstraction- and,
thus, variable-free. In doing so, we have replaced the variable
binding of the $\lambda$ operator with the variable \emph{routing} of
these basis combinators. See~\cite{liang10programs} for a more
detailed discussion of this routing interpretation.

This is very convenient for program synthesis: since every expression
is the application of one expression to another -- with this recursion
bottoming out at the primitive combinators -- each program is a binary
tree. Most importantly, any subtree is itself a well-formed
expression; this is not the case in the lambda calculus, since
abstraction introduces long range dependencies between the $\lambda$
operator and the variables to which that operator refers. In the
lambda calculus, then, a subtree might have free variables, not bound
to any enclosing $\lambda$.

@@

We can extend this representation with a simple polymorphic type
system~\cite{Pierce_2002}. In this representation, a type is either a
type primitive (e.g. reals, integers, booleans, etc.), a type variable
(e.g. $\sigma$), or a function type $ t_1 \rightarrow t_2$ of
functions from source type $t_1$ to target type $t_2$. Thus, this
combinator can be represented as a binary
tree~(Figure~\ref{fig:clbintree}) whose leaves are typed primitive
combinators and whose interior nodes represent typed applications of
combinators.

\begin{figure}
  \centering
  \begin{subfigure}[Before]{0.5\linewidth}
    \begin{tikzpicture}[scale=.9,
        transform shape, level/.style={sibling distance=10mm/#1, font=\scriptsize}
        ]
      \node [font=\scriptsize] (top) {$Int$}
      child {node [above=.3cm, text width=2cm] (a) {$(Int \rightarrow Int) \rightarrow Int 
          \rightarrow Int$}
        child {node [left=.5cm, below=1cm, text width=3cm] 
          {$S : \\(t_2 \rightarrow t_1 \rightarrow t_0)  \rightarrow (t_2 \rightarrow t_1) 
            \rightarrow \\ t_2 \rightarrow t_0$}}
        child {node [right=0cm, text width=2cm] {$*: Int \rightarrow Int \rightarrow Int$}}
      }
      child {node [below right=.3cm and 1cm] (b) {$Int \rightarrow Int$}
        child {node {$I: t_0 \rightarrow t_0 \rightarrow t_0$}}
      };
    \end{tikzpicture}
    \label{fig:clbintree}
  \end{subfigure}
  \begin{subfigure}[After]{0.4\linewidth}
    \begin{tikzpicture}[scale=.7, transform shape, 
        level/.style={sibling distance=10mm/#1},
        every edge/.style={draw, thin}
      ]
      \node [circle,draw] (a) {}
      child {node [left=0cm] (b) {$f_1$}}
      child {node [left=.2cm] (c) {\dots} edge from parent[draw=none] }
      child {node [left=.4cm] (e) {$f_n$}}
      child {node [circle,draw, left=1cm] (d) {} edge from parent[very thick]
        child {node [circle, draw, left=.25cm] (l) {} edge from parent[draw=none]
          child {node [left=.5cm] (b) {$f_1$}} 
          child {node [left=.3cm] (c) {\dots} edge from parent[draw=none] }
          child {node [left=.3cm] (e) {$f_n$} edge from parent[very thick]}
          child {node [circle, draw, left=.25cm] (and2) {}
            child {node [circle, draw, left=.25cm](l2) {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
            child {node [circle, draw] (r2)  {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
          }
        }
        child {node [circle, draw, right=.25cm] (r) {} edge from parent[draw=none]
          child {node [circle, draw, right=.25cm] (and3) {}
            child {node [circle, draw ](l3) {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
            child {node [circle, draw, right=.2cm ] (r3)  {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
          }
          child {node [right=.3cm] (b) {$f_1$} edge from parent[very thick]} 
          child {node [right=.4cm] (c) {\dots} edge from parent[draw=none] }
          child {node [right=.6cm] (e) {$f_n$}}
        }
      };

@@

\subsection{A stochastic grammar over programs.}
\label{sec:stochgrammar}
A stochastic grammar over programs represents a distribution over
expressions in $\mathcal{L}$. We associate with each primitive
combinator $c_i$ in $\mathcal{L}$ a probability $p_i$,
$0~\leq~p_i~\leq~1, \sum_i~p_i~=~1 $.

We define the probability of an expression $e$ consisting of primitive
combinators $c_1, \dots, c_N$ to be $p(e)~=~\prod_{n=1}^N p(c_n |
t(c_n))$ where $t(c_n)$ is the requested type when $c_n$ was used and
\[
p(c_i | t(c_n)) =    \frac{1}{4}\frac{p_i}{\sum_{c_j :
    type(c_j) \text{ unifies with } t(c_n)} p_j}.
\]
That is, $p(c_i | t(c_n)$ is the probability of sampling $c_i$ from
the multinomial distribution defined by the $p_i$'s, conditioned on
the requesting type, and then divided by 4. The constant factor of
$1/4$ is introduced as a simple way to ensure that the sum over the
probabilities of all expressions converges to finite value (so that
the distribution over expressions is proper).\footnote{The probability
  mass $M_d$ of all expressions of depth less than or equal to $d$ can
  be written with the recurrence relation $M_d \leq M_d^2 + M_1$. This
  has an asymptotic fixed point if $x = x^2 + M_1$ has a solution,
  which is true if $M_1 \leq 1/4$.}

>>>>

\usepackage{ijcai13}
\usepackage{times}
\usepackage{amsmath} 
\usepackage{latexsym} 
\usepackage{ dsfont }
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{float}
\usepackage{circuitikz}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{alltt}
\usepackage{dashrule}
\usepackage{array}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usetikzlibrary{arrows,automata, positioning, patterns,backgrounds}

\newtheorem{definition}{Definition}
\DeclareMathOperator*{\argmax}{arg\,max} 
\DeclareMathOperator*{\argmin}{arg\,min}

@@

\author{Eyal Dechter \\
MIT\\
USA \\
edechter@mit.edu
\And
Jon Malmaud \\
MIT \\ 
USA \\
malmaud@mit.edu
\And 
Ryan P.  Adams \\
Harvard University\\
USA \\
rpa@seas.harvard.edu
\And
Joshua B. Tenenbaum \\
MIT\\
USA \\
jbt@mit.edu
}

@@

\section{Introduction}
Many of the problems we want to solve in AI are best cast as search
problems. There are many search algorithms, but they typically exploit
local or topological structure in the search space. For example,
convex optimization and gradient based search methods~\cite{} take
advantage of local smoothness to move in good directions. Local search
algorithms move from one candidate solution to a neighboring
one\cite{}. Heuristic search techniques take advantage of topological
structure to prune off large parts of the solution space~\cite{}.

However, many of the problems we want solve -- discovering scientific
theories, developing new technologies, figuring out how to bake -- do
not come with such a structured search space. This is particularly
true of problems, as these are, whose solutions can be naturally
framed as programs. Here we ask how to extract structure from the
space of programs by attempting to solve multiple related problems in
a given domain at once.

We take as inspiration the programmer who -- having reused several
motifs, programming patterns, or subroutines to solve different but
similar simple tasks -- creates a function library to help manage the
complexity of larger projects. Our algorithm builds a library
of reusable program components and a distribution over these,
effectively learning from simple tasks a search space structure which
enables it to solve more complex tasks.

In this work, we try to see how far this idea of reusable function
definition can get us. Our question is, Can an AI system discover the
structure latent in the solutions to multiple related problems and
thereby bootstrap effective exploration of the search space? Can
discovering modular program components transform a problem from one in
which search is intractable into a problem in which it becomes
increasingly feasible with experience?

Concretely, our algorithm works as follows. It begins with a small
library of program primitives and a distribution over programs
synthesized from this library. In each iteration, it generates
programs in order of highest probability according to the library
learned in the previous iteration. From the discovered solutions, it
finds new reusable functions that compress these solutions, and it
re-estimates the distribution over programs by weighting these new
functions according to their frequencies.

We show that this iterative approach allows us to achieve very high
performance in tasks where just searching according to the initial
distribution over program primitives clearly fails.

\section{Related Work}
The idea of discovering and using reusable subcomponents is an old one
in AI. Rendell's 1985 paper~\cite{DBLP:conf/ijcai/Rendell85},
``Substantial Constructive Induction Using Layered Information
Compression: Tractable Feature Formation in Search,'' -- which could
have been the name of this paper -- did not use programs as its
solution space representation. Within the Inductive Logic Programming
literature, researchers have explored ``Constructive Induction,''
generating new and reusable logic programming
terms~\cite{DBLP:conf/ijcai/Muggleton87}. We believe that applying
constructive induction to a compositional functional programming
paradigm will provide even greater benefits.

Dicovering useful subroutines has been explored in Genetic
Programming as
well~\cite{DBLP:conf/foga/Koza92}~\cite{DBLP:conf/ices/KozaBAK96}.

The importance of a curriculum has also been explored previously,
though generally the curriculum has been explicitly given rather than
discovered. Quinlan and Cameron-Jones~\cite{Quinlan93foil:a} showed
that a logic program learning algorithm could learn sequentially more
difficult list programs if the algorithm was provided with its own
solutions to previously solved simple problems. More recently, several
researchers have observed that presenting problems to machine learning
algorithms in order of difficulty results in better
performance~\cite{Krueger2009380}~\cite{DBLP:conf/icml/BengioLCW09}.

Our use of a stochastic grammar over combinatory logic is based on
Liang et al.~\cite{liang10programs}. In that work, a non-parametric
distribution was placed over combinatory logic expressions, and an
MCMC algorithm was developed to perform joint inference over these
expressions and the underlying stochastic grammar. Our approach is
inspired by this generative model, but avoids the difficulties
associated with a stochastic local search over programs (the same
difficulties faced by genetic programming), especially when the
likelihood function (or the fitness function) are not of much help.

%% AI and machine learning algorithms commonly take advantage of the
%% relationship between a performance metric (error, reward, fitness,
%% etc.) and the local or topological structure of the solution
%% space. Whether in discrete or continuous domains, the key is to design
%% such a relationship and write a algorithm that either moves in good
%% directions or prunes off large parts of the solution space. Whether or
%% not such a relationship exists depends on the representation, and much
%% work has gone into finding good representations for problem
%% solving. Recent work, especially in machine learning, has focused on
%% ``feature learning,'' with the explicit aim of discovered good
%% representational building blocks in vector-valued solutions spaces.

%% Many of the most natural solution representations for intelligent
%% agents, however, are not naturally expressed as vector-valued
%% spaces. Consider the problems of discovering scientific theories,
%% inventing new technologies, communicating with others, or executing a
%% complex motor routine (such as dancing or playing an instrument). The
%% solutions to these problems are most naturally expressed as programs,
%% that is, as the composition of modular components whose semantics are
%% determined by the actions of one component on another.

%% How can we learn good solution representations for these kinds of
%% tasks without building in the kind of domain knowledge that gives rise
%% to a useful relationship between representation and performance? How
%% can we get traction on a set of problems when small tweaks to
%% solutions give rise to wildly varying performances in an unpredictable
%% manner, or when, for most problems, most of the solutions fail in a
%% uniformly bad way?

%% A popular idea is that good representations are those that encode the
%% data space in an efficient way, that is, that enable compression of
%% the data~\cite{}. For program representations, this most naturally
%% corresponds to defining reusable components so that the total number
%% of unique expressions for a set of solutions is minimized (i.e., this
%% is directly akin to the subroutine libraries programmers use to manage
%% the complexity of their code).

%% In this work, we try to see how far reusable subroutine definition can
%% get us on its own. Our question is, Can discovering modular components
%% transform a problem from one in which brute-force search will not work
%% to a problem in which it does?

%% To emphasize an impoverished relationship between reward signals and
%% program representations, we make our tasks all-or-none: a solution
%% either solves the problem or it does not, and partial solutions do not
%% get any credit.

%% The intuitive approach to applying this idea is simple: if solutions
%% to problems in the domain share common structure, and there exist
%% problems whose solutions are simple enough to find by brute force,
%% then we should be able to extract a distribution over solution
%% subcomponents from these simple cases that allows us to run subsequent
%% searches more efficiently.

%% More concretely, our algorithm works as follows: each iteration, it
%% first generates programs in order of highest probability according to
%% a subroutine distribution learned in the previous iteration. From the
%% discovered solutions, it finds new reusable subroutines that compress
%% these solutions and estimates their probabilities.

%% We show that this iterative approach allows us to achieve very high
%% performance in tasks where just searching according to the initial
%% distribution over program primitives clearly fails.

%%  \hline

%% AI and machine learning algorithms commonly take advantage of the idea
%% that local structure in the solution space is related to local
%% structure in the outcome. In continuous domains, learning algorithms
%% algorithms use error gradients to guide search. In discrete domains,
%% whether in heuristic or local search paradigms, knowledge about the
%% solution space topology and its relation to reward and cost metrics
%% enables pruning, reasoning, and moving in the solution space more
%% efficiently. However, choosing the representations that have
%% beneficial structural properties requires a lot of general and domain
%% knowledge, and this what researchers do and what they cannot leave up
%% to their algorithms.

%% What is a good strategy for an agent when its representation of the
%% solution space is not smooth with respect to the outcome? Many hard
%% learning problems have this property, especially those in which the
%% solution space is most naturally represented as a set of
%% programs. Consider the problems of discovering scientific theories,
%% inventing new technologies, learning to cook, or learning to
%% communicate. These are problems in which 1) solutions are best
%% described as programs (i.e. compositions of modular subroutines at
%% various levels of abstraction), 2) most of the solution space is bad
%% in a relatively uninformative way (i.e. it just doesn't work), and
%% thus 3) outcomes are not informative about how you are doing unless
%% you are already quite close to a solution.

%% Intuitively, the way to solve such problems is to find solutions to
%% simple cases in the domain and extract conceptual building blocks that
%% can be composed to scale up to complex cases. With experience, the
%% agent may be able to discover local structure in the solution space,
%% and utilize more traditional AI and machine learning techniques. In
%% this work, we will be focusing on the first phase of this learning
%% problem.

%% %% Of course, every learner

%% %% Choosing the wrong representation, in
%% %% which local structure is not apparent, is often the difference between
%% %% a successful learner and an unsuccessful one. In many respects, this
%% %% may be the kind of situation that adult humans face when confronting
%% %% new domains: we have a lot of general knowledge of successful
%% %% representations, useful strategies, etc.

%% %% But suppose that you have very little knowledge of a domain. Your
%% %% intial representation of the domain is poor: ``nearby'' points in the
%% %% solutions space ellicit wildly different outcomes in unpredictable
%% %% ways. Or perhaps nearby points ellicit the same outcome and its not a
%% %% good one, so you don't know which way to go. Many learning problems,
%% %% especially those whose solutions are best described as programs, have
%% %% these qualities: imagine learning to synthesize new sentences

%% %% teness. We choose the representations our learners use
%% %% and these representations generally Imagine you are trying to learn to
%% %% bake via trial and error. You know what kind of ingredients are used
%% %% in baking (flour, water, butter, sugar, yeast, baking soda, etc.) and
%% %% you know a bit about how to combine these ingredient (mix, whisk,
%% %% etc.) and how to apply heat (oven). But baking is a notoriously
%% %% fragile process: ratios of ingredients, the order of mixing, cooking
%% %% temperatures, resting times, all make a huge difference, and the
%% %% outcome of the baking process is mostly all or none. Either the
%% %% outcome is what you expected or quite unlike it. Even if you do create
%% %% something similar to your intent, it is difficult to assign blame for
%% %% the outcome to any one step or decision along the way.

%% %% Many difficult learning problems have this structure -- learning
%% %% language, developing technologies and scientific theories, learning to
%% %% play instruments or execute other complex motor patterns. These are
%% %% problems in which 1) there is a large set of similar tasks; 2) their
%% %% solutions are best described as programs; 3) the solution space is
%% %% not smooth with respect to the outcome; 4) and the outcome is generally
%% %% uninformative unless you are already pretty close to a solution.

%% %% Of course, knowledge is not acquired wholly by trial and error. But in
%% %% many domains, there is at least initially such a period of ignorance,
%% %% and the question we are trying to tackle here is how you ever get off
%% %% the ground.

%% I need help here.

%% %% Let a \emph{task} be a function $t : \mathcal{L} \rightarrow \{0,
%% %% 1\}$, where $\mathcal{L}$ is the set of all expressions in our
%% %% representation language. We say that $e \in \mathcal{L}$ \emph{solves}
%% %% task $t$ if $t(e) = 1$. Our goal is to solve as many tasks as
%% %% possible, as efficiently as possible. That is, suppose that there is a
%% %% distribution over tasks $t \sim  p_t(\cdot)$, we would like an ordering $O
%% %% = e_1, \dots $ on expressions in $\mathcal{L}$ such that if $p_t(t_1)
%% %% > p_t(t_2)$ then the first expression in $O$ to solve $t_1$ comes
%% %% before the first expression in $O$ to solve $t_2$.

%% %% We motivate a solution to this problem by considering a heirarchical
%% %% Bayesian formulation of the task distribution. Suppose we believe that
%% %% there is some underlying structure to $p_t(\cdot)$ that is explained by
%% %% the fact that for each task $t$ there is a latent expressions $e \in
%% %% \mathcal{L}$ drawn from a stochastic grammar $\mathcal{G}$ and that

%% %% \begin{align}
%% %% e &\sim p_\mathcal{L}(\cdot) = \mathcal{G}\\
%% %% t | e &\sim \text{U}\{ \text{tasks solved by }e\}.
%% %% \end{align}

%% %% Then $p_t(t) = \sum_{e_j \in \mathcal{L} \text{ s.t. } t(e_j) = 1}
%% %% p_\mathcal{L}(e_j)$. If we make the assumption that for the set $E$ of
%% %% expressions in this sum, there is some expression $e^*$ such that
%% %% $p_\mathcal{L}(e^*) >> p_\mathcal{L}(e), e \neq e^*$, then we have, approximately, 
%% %% \[ p_t(t) \propto} p_\mathcal{L}(e^*). \]

%% %% This suggests a way to find the ordering $O$: infer a stochastic
%% %% grammar $G$, under the assumption that $G$ favors one solution to a
%% %% given task over any other, and order the elements of $\mathcal{L}$
%% %% according to $G$.

%% %% Thus, our basic procedure will be to search for a single ``good''
%% %% solution for as many tasks as we can, and to use those solutions to
%% %% estimate a stochastic grammar. Our search will be performed in the
%% %% order of the most recently inferred grammar. So as our ordering
%% %% becomes more efficient and we can solve more tasks, we learn better
%% %% stochastic grammars.

%% %% Formally then, our goal is to solve the following minimal feedback
%% %% multi-task program induction problem: suppose we are given set of
%% %% \emph{tasks} $T=\{t_k\}_{k=1}^K$ where each task is a function $t_i :
%% %% \mathcal{L} \rightarrow \{0, 1\}$, where $\mathcal{L}$ is the set of
%% %% programs in our language.  We say that $ p \in \mathcal{L}$
%% %% \emph{solves} $t_i$ if $ t_i(p) = 1$. Our goal is to find a set of
%% %% solutions $X = \{x_k\}_{k=1}^k \subset \mathcal{L}$ such that

%% %% \[
%% %% X = \argmax_{x_1, \dots, x_K \in L} \sum_k t_k(x_k).
%% %% \]

%% Based on the motivation set out above, we present an iterative
%% strategy. Given a stochastic grammar $\mathcal{G}$ which defines a
%% distribution over the primitives from $\mathcal{L}$ and a integer
%% \emph{frontier size} $N$, we perform two steps in each iteration.  In
%% the first step, we explore the $N$ highest probability programs
%% according to $\mathcal{G}$. In the second step, we compress the
%% programs that solve any of our tasks, and we use that compression to
%% define a new grammar. Repeating these two steps, we effectively
%% reparameterize our search space to include programs that are more likely
%% to be solutions to our tasks.

The remainder of this paper will include two major sections.  We will
first describe the representation language we chose and the details of
the exploration and compression steps of the algorithm. We will then
present results from two experiments involving the domains of
symbolic regression and boolean circuit learning.

@@

\section{Problem statement and terminology}
Our goal is to solve the following minimal feedback multi-task program
induction problem: suppose we are given set of \emph{tasks}
$T=\{t_k\}_{k=1}^K$ where each task is a function $t_i : \mathcal{L}
\rightarrow \{0, 1\}$ and where $\mathcal{L}$ is the set of programs
in our language.  We say that $ p \in \mathcal{L}$ \emph{solves} $t_i$
if $ t_i(p) = 1$. Our goal is to solve as many of the tasks in $T$ as
we can.

We will use the following terminology. A \emph{frontier} will be the
finite set of programs that the algorithm is considering at any one
time. We will say that a task is \emph{hit} by the frontier if there
is a program in the frontier that solves it.

@@

Following~\cite{liang10programs} and~\cite{Briggs:2008}, we use a
simply typed combinatory logic -- a variable-free subset of the
polymorphic simply-typed lambda calculus~\cite{Pierce_2002} -- as our
program representation language. In short, the combinatory logic
introduces a \emph{basis} of several primitive \emph{combinators} such
that any function in the lambda calculus can be alternatively written
as applications of the basis combinators. Some common basis
combinators are defined as follows:
\begin{align}
I\; x &\rightarrow  x \\
S\; f\; g\; x\; &\rightarrow (f\; x)\; (g\; x) \\ 
B\; f\; g\; x\; &\rightarrow (f\; x)\; g \\ 
C\; f\; g\; x\; &\rightarrow f\; (g\; x) \\ 
\end{align}

The basis combinators are theoretically enough to express any Turing
computable function, but we will assume that our language has a
variety of primitives and, together with the basis combinators, we
call these the primitive combinators. Note that by default we consider
all primitives to be in their ``curried'' form (e.g. a function like
``+'' adds two numbers $x$ and $y$ by being first applied to $x$,
returning the function $(+\;x)$ and then being applied to $y$,
returning the object $(+\;x)\;y$).

The basis combinators can themselves been expressed in the lambda
calculus (with definitions following directly from the equations
above). The lambda calculus has two basic operations -- application
and abstraction -- but in using the combinatory logic we sequester
uses of the abstraction operation inside the combinators. In doing so,
we have replaced the variable binding of the $\lambda$ operator with
the variable \emph{routing} of these basis combinators. Our
representation thus becomes variable-free. See Liang et
al.~\cite{liang10programs} for a more detailed discussion of this
routing interpretation.

Using combinatory logic is very convenient for program synthesis:
since every expression is the application of one expression to another
-- with this recursion bottoming out at the primitive combinators --
each program is a binary tree. Most importantly, any subtree is itself
a well-formed expression; this is not the case in the lambda calculus,
since abstraction introduces long range dependencies between the
$\lambda$ operator and the variables to which that operator refers. In
the lambda calculus, then, a subtree might have free variables, not
bound to any enclosing $\lambda$.

@@

We can extend this representation with a simple polymorphic type
system~\cite{Pierce_2002}. In this representation, a type is either a
type primitive (e.g. reals, integers, booleans, etc.), a type variable
(e.g. $\sigma$), or a function type $ t_1 \rightarrow t_2$ of
functions from source type $t_1$ to target type $t_2$. Any combinator
can be represented as a binary tree~(Figure~\ref{fig:clbintree}) whose
leaves are typed primitive combinators and whose interior nodes
represent typed applications of combinators (we annotate interior
nodes with their types).

\begin{figure}
%%  \centering
  \begin{subfigure}[]{0.45\linewidth}
    \begin{tikzpicture}[scale=.8,
        transform shape, level/.style={sibling distance=5mm/#1, font=\scriptsize}
        ]
      \node [font=\scriptsize] (top) {$\mathbb{R}$}
      child {node [above=.3cm, left=.1cm, text width=1.8cm] (a) {$(\mathbb{R} \rightarrow \mathbb{R}) \rightarrow \mathbb{R} 
          \rightarrow \mathbb{R}$}
        child {node [left=.5cm, below=1cm, text width=3.0cm] 
          {$ \mathbf{S} : (t_2 \rightarrow t_1 \rightarrow t_0)  \rightarrow (t_2 \rightarrow t_1) 
            \rightarrow \\ t_2 \rightarrow t_0$}}
        child {node [right=0cm, text width=2cm] {$ \mathbf{*}: \mathbb{R} \rightarrow \mathbb{R} \rightarrow \mathbb{R}$}}
      }
      child {node [below right=-.5cm and .2cm] (b) {$\mathbb{R} \rightarrow \mathbb{R}$}
        child {node [above=.5cm] {$ \mathbf{I}: t_0 \rightarrow t_0 \rightarrow t_0$}}
      }
    \end{tikzpicture}
    \caption{}
    \label{fig:clbintree}
  \end{subfigure}
  \begin{subfigure}[R]{0.5\linewidth}
    \begin{tikzpicture}[scale=.7, transform shape, 
        level/.style={sibling distance=10mm/#1},
        every edge/.style={draw, thin}
      ]
      \node [circle,draw] (a) {}
      child {node [left=0cm] (b) {$f_1$}}
      child {node [left=.2cm] (c) {\dots} edge from parent[draw=none] }
      child {node [left=.4cm] (e) {$f_n$}}
      child {node [circle,draw, left=1cm] (d) {} edge from parent[very thick]
        child {node [circle, draw, left=.25cm] (l) {} edge from parent[draw=none]
          child {node [left=.5cm] (b) {$f_1$}} 
          child {node [left=.3cm] (c) {\dots} edge from parent[draw=none] }
          child {node [left=.3cm] (e) {$f_n$} edge from parent[very thick]}
          child {node [circle, draw, left=.25cm] (and2) {}
            child {node [circle, draw, left=.25cm](l2) {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
            child {node [circle, draw] (r2)  {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
          }
        }
        child {node [circle, draw, right=.25cm] (r) {} edge from parent[draw=none]
          child {node [circle, draw, right=.25cm] (and3) {}
            child {node [circle, draw ](l3) {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
            child {node [circle, draw, right=.2cm ] (r3)  {} edge from parent[draw=none] 
              child {node [above=.5cm] {\vdots} edge from parent[draw=none]}
            }
          }
          child {node [right=.3cm] (b) {$f_1$} edge from parent[very thick]} 
          child {node [right=.4cm] (c) {\dots} edge from parent[draw=none] }
          child {node [right=.6cm] (e) {$f_n$}}
        }
      };

@@

\subsection{A stochastic grammar over programs.}
\label{sec:stochgrammar}
We would like a way to describe how likely we believe a program to be
given the probabilities we ascribe to its components. Recall that
$\mathcal{L}$ is the set of all programs.  A stochastic grammar over
programs represents a distribution over expressions in
$\mathcal{L}$. We associate with each primitive combinator $c_i$ in
$\mathcal{L}$ a probability $p_i$, $0~\leq~p_i~\leq~1, \sum_i~p_i~=~1
$.

The probability of using a primitive combinator somewhere in a program
should depend on which other combinators could have been used in its
place. This, in turn, depends on the type at that particular location
in the program tree; we call this type the \emph{requesting} type. For
example, suppose that we have the primitive $(+1)$ of type $Int
\rightarrow Int$ and the primitive $NOT$ of type $Bool \rightarrow
Bool$. If the requesting type is $\sigma \rightarrow Int$, then we can
only use $(+1)$, but if the requesting type is $\sigma \rightarrow
\sigma$ then we could use either $(+1)$ or $NOT$. Let $C = c_1, \dots,
c_n$ be the set of primitive combinators in our grammar; we define
$C_t$ for any type $t$ to be the set of combinators in $C$ that are
consistent with $t$, that is, whose types \emph{unify} with t (we use
unification according to the standard type theory
presentation~\ref{Pierce_2002}).
